RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2014-12-02 manuscripts/manuscript.Rmd
Clarify lack of OTU block in ape::phylo & explain merging character blocks into data frame with get_characters as suggested by @hlapp in #101

Conceptually, a NeXML document has the following components: (1) phylogeny
data, (2) character or trait data in matrix form, (3) operational
taxonomic units (OTUs), and (4) metadata. When reading and writing NeXML,
`RNeXML` aims to map these to and from, respectively, their most
natural representations in R. These are the `phylo` object from the
`ape` package [@Paradis_2004], R's native `data.frame` list structure for
phylogenies, character matrices, and metadata, respectively.
<!-- compared to the preceding enumeration, this is missing where the OTUs go. HL -->

@@

To allow strict validation of data types, the NeXML schema requires
continuous and discrete character data to be separated into different
"characters" blocks. Since the `data.frame` data structure in R has no
such constraints, the `get_characters()` method will combine such
blocks as separate columns into a single `data.frame` object, provided
they correspond to the same taxa. Otherwise, a list of `data.frame`s
is returned, with list elements corresponding to characters
blocks. Similar to the distinction between `get_trees()` and
`get_trees_list()`, there is also a method `get_characters_list()`,
which always returns a list of `data.frame`s.

>>>>

Conceptually, a NeXML document has the following components: (1) phylogeny
data, (2) character or trait data in matrix form, (3) operational
taxonomic units (OTUs), and (4) metadata. When reading and writing NeXML,
`RNeXML` aims to map these to and from, respectively, their most
natural representations in R. These are the `phylo` object from the
`ape` package [@Paradis_2004], R's native `data.frame` list structure for
phylogenies, character matrices, and metadata, respectively. Note that
unlike NeXML or NEXUS, the `phylo` object stores taxonomic information
only as tip labels and not as a separate block of data.

@@

Observe that `RNeXML` has automatically merged the different NeXML character
blocks, (representing continuous and discrete characters respectively) 
into a single data frame with appropriate column classes. 
To allow strict validation of data types, the NeXML schema requires
continuous and discrete character data to be separated into different
"characters" blocks. Since the `data.frame` data structure in R has no
such constraints, the `get_characters()` method will combine such
blocks as separate columns into a single `data.frame` object, provided
they correspond to the same taxa. Otherwise, a list of `data.frame`s
is returned, with list elements corresponding to characters
blocks. Similar to the distinction between `get_trees()` and
`get_trees_list()`, there is also a method `get_characters_list()`,
which always returns a list of `data.frame`s, one for each
character block.
